### 2021-03-08 README.Rmd
updated docs #22

>>>>

## Package data

### Package data objects

Unexported objects `ped1kg` and `ped1kg_unrel` contain [pedigree data obtained from 1000 Genomes FTP](http://ftp.1000genomes.ebi.ac.uk/vol1/ftp/release/20130502/integrated_call_samples_v3.20200731.ALL.ped) for all 2504 individuals with sequencing data, and all individuals unrelated to anyone else in the data (momid, dadid, siblings, second_order, and third_order all =0). This code gets all the unique IDs for anyone listed as a mother, father, sibling, child, second order, or third order relative, _who also has sequencing data_, and removes any _individual ID_ who matches one of these relatives. See [data-raw/generate_sysdata.R](data-raw/generate_sysdata.R), specifically the code that begins with the `relatives <-` assignment.

```{r}
skater:::ped1kg
skater:::ped1kg_unrel
```

There are **`r nrow(skater:::ped1kg)-nrow(skater:::ped1kg_unrel)`** samples who have known relatives in the 1000 Genomes ped file (included in `ped1kg` but not in `ped1kg_unrel`):

```{r}
skater:::ped1kg %>% 
  dplyr::anti_join(skater:::ped1kg_unrel) %>% 
  dplyr::select(fid, id, population, second_order:other_comments) %>% 
  knitr::kable()
```

### Text files with unrelated founder sample IDs

After creating the `ped1kg_unrel` object in [data-raw/generate_sysdata.R](data-raw/generate_sysdata.R) as described above, the code in [data-raw/write-sampleids-1000g-unrelated.R](data-raw/write-sampleids-1000g-unrelated.R) then writes one sample ID per line for each of those unrelated samples to `inst/extdata`. These can be used to pass to `bcftools view --samples-file <POP.txt>`.

The directory containing these files can be found on any system where skater is installed:

```{r}
system.file("extdata", "sampleids-1000g-unrelated", package="skater", mustWork=TRUE)
```

The table below shows how many unrelated founders are available in any given population:

```{r}
system.file("extdata", "sampleids-1000g-unrelated", package="skater", mustWork=TRUE) %>% 
  list.files(full.names=TRUE) %>% 
  rlang::set_names(basename) %>% 
  purrr::map_int(~length(readr::read_lines(.))) %>% 
  tibble::enframe() %>% 
  knitr::kable()
```
